---
title: "Calibrated model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calibrated model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(misc)
```

# Load mtcars dataset
```{r}
data(mtcars)
```

# Split data into training and test sets
```{r}
set.seed(123)
train_idx <- sample(nrow(mtcars), size = floor(0.7 * nrow(mtcars)))
train_data <- mtcars[train_idx, ]
test_data <- mtcars[-train_idx, ]

# Fit regular linear model
```{r}
lm_model <- lm(mpg ~ ., data = train_data)
print(summary(lm_model))
print(confint(lm_model))
```

# Fit calibrated model 
```{r}
(calib_model <- misc::calibmodel(X = as.matrix(train_data[,-1]), y = train_data$mpg))
print(summary(calib_model))
print(confint(calib_model))
```

# Make predictions on test set
```{r eval=TRUE}
lm_pred <- predict(lm_model, newdata = test_data, interval = "prediction")
calib_pred <- predict(calib_model, newdata = as.matrix(test_data[,-1]), interval = "prediction")
```


# Compare predictions
```{r eval=TRUE, fig.width=7.5}
results <- data.frame(
  Actual = test_data$mpg,
  LM_Pred = lm_pred[,"fit"],
  LM_Lower = lm_pred[,"lwr"],
  LM_Upper = lm_pred[,"upr"],
  Calib_Pred = calib_pred[,"fit"],
  Calib_Lower = calib_pred[,"lwr"], 
  Calib_Upper = calib_pred[,"upr"]
)

# Print results
print("Prediction Intervals Comparison:")
print(results)

# Calculate coverage (% of actual values within prediction intervals)
lm_coverage <- mean(test_data$mpg >= results$LM_Lower & 
                   test_data$mpg <= results$LM_Upper)
calib_coverage <- mean(test_data$mpg >= results$Calib_Lower & 
                      test_data$mpg <= results$Calib_Upper)

print(sprintf("\nPrediction interval coverage:"))
print(sprintf("Linear Model: %.1f%%", 100 * lm_coverage))
print(sprintf("Calibrated Model: %.1f%%", 100 * calib_coverage))

# Set common y-axis limits for both plots
y_limits <- range(c(results$LM_Lower, results$LM_Upper,
                   results$Calib_Lower, results$Calib_Upper))

# Plot prediction intervals
par(mfrow=c(1,2))

# Linear Model Plot
plot(results$Actual, results$LM_Pred, 
     main="Linear Model Predictions",
     xlab="Actual MPG", ylab="Predicted MPG",
     ylim=y_limits)
# Add shaded prediction intervals
x_ordered <- order(results$Actual)
polygon(c(results$Actual[x_ordered], rev(results$Actual[x_ordered])),
        c(results$LM_Lower[x_ordered], rev(results$LM_Upper[x_ordered])),
        col=rgb(0, 0, 1, 0.2), border=NA)
points(results$Actual, results$LM_Pred)  # Replot points over shading
abline(0, 1, col="red", lty=2)  # Add diagonal line

# Calibrated Model Plot
plot(results$Actual, results$Calib_Pred,
     main="Calibrated Model Predictions",
     xlab="Actual MPG", ylab="Predicted MPG",
     ylim=y_limits)
# Add shaded prediction intervals
polygon(c(results$Actual[x_ordered], rev(results$Actual[x_ordered])),
        c(results$Calib_Lower[x_ordered], rev(results$Calib_Upper[x_ordered])),
        col=rgb(0, 0, 1, 0.2), border=NA)
points(results$Actual, results$Calib_Pred)  # Replot points over shading
abline(0, 1, col="red", lty=2)  # Add diagonal line
```

